---
title: "HW 2: From correlation to linear mixed-effect models. Assignment sheet"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lme4)
library(vcd)
library(ggplot2)
```

## Задание 1

### 1.0

``` {r}
shva <- read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/duryagin_ReductionRussian.txt', sep='\t')
```

### 1.1

``` {r}
ggplot(data=shva, aes(f2, f1)) + geom_point(aes(colour=factor(shva$vowel)), show.legend = F) + scale_y_reverse() + scale_x_reverse() + labs(title='f2 and f1 of the reduced and stressed vowels', caption="Data from Duryagin 2018")
```

### 1.2 Plot the boxplots of f1 and f2 for each vowel using ggplot().

``` {r}
ggplot(data=shva, aes(vowel, f1)) + geom_boxplot(aes(fill=factor(shva$vowel)), show.legend = F) + coord_flip() + labs(title='f1 distribution for each vowel', caption="Data from Duryagin 2018")
ggplot(data=shva, aes(vowel, f2)) + geom_boxplot(aes(fill=factor(shva$vowel)), show.legend = F) + coord_flip() + labs(title='f2 distribution for each vowel', caption="Data from Duryagin 2018")
```

### 1.3 Which f1 can be considered outliers in a vowel?

3 точки из y, одна точка из A, три точки из а

### 1.4 Calculate Pearson’s correlation of f1 and f2 (all data)

``` {r}
cor(shva$f1, shva$f2, method='pearson')
```

### 1.5 Calculate Pearson’s correlation of f1 and f2 for each vowel 

``` {r}
a <- shva[shva$vowel=='a',] 
y <- shva[shva$vowel=='y',]
A <- shva[shva$vowel=='A',]
print(cor(a$f1, a$f2))
print(cor(y$f1, y$f2))
print(cor(A$f1, A$f2))
```
## 1.6 Use the linear regression model to predict f2 by f1.

``` {r}
m <- lm(f2~f1, data=shva)
```
#### 1.6.1 Provide the result regression formula

``` {r}
coefficients(m)
```
> 1639.7021488 - 0.4287498 f1

#### 1.6.2 Provide the adjusted R2
``` {r}
summary(m)
```

> R-squared:  0.3361

#### 1.6.3 Add the regression line in scatterplot 1.1

``` {r}
for_the_picture <- lm(f1~f2, data=shva)
ggplot(data=shva, aes(f2, f1)) + 
  geom_point(aes(colour=factor(shva$vowel)), show.legend = F) +
  geom_line(aes(f2, predict(for_the_picture))) +
  scale_y_reverse() + 
  scale_x_reverse() + 
  labs(title='f2 and f1 of the reduced and stressed vowels', caption="Data from Duryagin 2018")
```

### 1.7 Use the mixed-efects model to predict f2 by f1 using vowel intercept as a random effect

``` {r}
m2 <- lmer(f2~f1+(1|vowel), data=shva)
```

#### 1.7.1 Provide the fixed effects formula

``` {r}
fixef(m2)
```

#### 1.7.2 Provide the variance for intercept argument for vowel random effects

``` {r}
summary(m2)
```

Variance:
> 6374

## Задание 2

### 2.0 Read the data from file to the variable elp.

``` {r}
elp <- read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/ELP.csv')
```

### 2.1 Which two variables have the highest Pearson’s correlaton value.

``` {r}
cor(elp$SUBTLWF, elp$Length, method='pearson')
cor(elp$Mean_RT, elp$SUBTLWF, method='pearson')
cor(elp$Mean_RT, elp$Length, method='pearson')
```

Ответ: `Mean_RT` ~ `Length`.



### 2.2 Group your data by parts of speech and make a scatterplot of SUBTLWF and Mean_RT.
```{r 2.2}
elp %>%
  ggplot(aes(SUBTLWF, Mean_RT)) +
  geom_point(aes(colour = Length)) +
  scale_color_continuous(low = "lightblue", high = "red") +
  facet_wrap(~POS, scales = "free_x") +
  scale_x_log10() +
  theme_bw() +
  labs(caption="data from (Balota et al. 2007)")
```

We've used `scale_color_continuous(low = "lightblue", high = "red")` as a parameter of `ggplot()`.

### 2.3 Use the linear regression model to predict `Mean_RT` by `log(SUBTLWF)` and `POS`.
m <- lm(Mean_RT ~ log(SUBTLWF) + POS, data=elp)
coefficients(m)

#### 2.3.1 Provide the result regression formula
```{r 2.3.1}

```

#### 2.3.2 Provide the adjusted R$^2$
```{r 2.3.2}
summary(m)$adj.r.squared
```

#### 2.3.3 Add the regression line in the scatterplot.
```{r 2.3.3}
mod <- lm(Mean_RT~log(SUBTLWF), data=elp)
model_prediction <- predict(mod)

elp %>%
  ggplot(aes(log(SUBTLWF), Mean_RT)) +
  geom_point(aes(colour = Length)) +
  scale_color_continuous(low = "lightblue", high = "red") +
  geom_line(aes(log(SUBTLWF), model_prediction), color='black') +
  theme_bw() +
  labs(caption="data from (Balota et al. 2007)")
```

### 2.4 Use the mixed-efects model to predict `Mean_RT` by `log(SUBTLWF)` using POS intercept as a random effect

m <- lmer(Mean_RT~log(SUBTLWF) + (1 | POS), data=elp)
summary(m)

fixef(m)

#### 2.4.1 Provide the fixed effects formula
```{r 2.4.1}

```

#### 2.4.2 Provide the variance for intercept argument for `POS` random effects
```{r 2.4.2}
414.4

```

#### 2.4.3 Add the regression line to the scatterplot
```{r 2.4.3}
m <- lmer(Mean_RT~log(SUBTLWF) + (1 | POS), data=elp)
elp$model_prediction <- predict(m)

elp %>% 
  ggplot(aes(log(SUBTLWF), Mean_RT, color=POS)) + 
  geom_point() + 
  facet_wrap(~POS) + 
  geom_line(aes(log(SUBTLWF), model_prediction), color='black') + 
  labs(caption='data from (Balota et al. 2007)') + 
  theme(legend.position='none')

```

## Задание 3

### 3.0

``` {r}
data <- read.csv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv', sep = ',')
```

### 3.1
Критерий χ2 плохо использовать, если хотя бы в одной из клеток ожидаемое значение меньше 5
``` {r}
table <- table(data$Aux, data$CeSynt) 
chisq.test(table)
```
получаем Warning: In chisq.test(table) : Chi-squared approximation may be incorrect

``` {r}
chisq.test(table)$expected
```


Понятно почему: в оюной из клеток значение 3.74. Если попробовать то же самое с остальными -- всё хорошо. Следовательно, ответ -- `Aux` ~ `CeSynt`.

``` {r}
fisher.test(table)
```


p-value < 2.2e-16. Следовательно, значима.

### 3.2 Test the hypothesis that Aux and EPTrans are not independent with the help of Pearson’s Chi-squared Test.

``` {r}
table <- table(data$Aux, data$EPTrans) 
chisq.test(table)
```
p-value = 0.0001553 -> yes, they are dependent.

### 3.3 Provide expected values for Pearson’s Chi-squared Test of Aux and EPTrans variables.
``` {r}
chisq.test(table)$expected
```

### 3.4 Calculate the odds ratio
``` {r}
fisher.test(table)$estimate
```
odds ratio = 2.601174

### 3.5 Calculate effect size for this test using Cramer’s V (phi).
``` {r}
library(questionr)
cramer.v(table)
```

### 3.6 Report the results of independence

We have found a significant association between variables `Aux` and `EPTrans` (p < 0.001). Effect size is small (Cramer's V = 0.1744882).


### 3.7 Visualize the distribution using mosaic plot.
``` {r}
library(vcd)
mosaic(~ Aux + CrSem + Country, data=data, shade=TRUE, legend=TRUE)
```

### 3.8

https://xkcd.com/882/